---
title: "Lab 2 Lightning Ridge Proposal"
author: "Luc R, Andrew B, Vish S, Arun N"
date: '2022-03-21'
output: html_document

---


# Installing all the packages
```{r load packages, echo=FALSE, message=FALSE}

install.packages('lmtest')
install.packages('patchwork')
install.packages('sandwich')
install.packages('forecast')
install.packages('corrplot')
install.packages('ggpubr')

library(gridExtra)
library(lattice)
library(ggpubr)
library(corrplot)
library(tidyverse)
library(magrittr)
library(ggplot2)
library(patchwork)
library(sandwich)
library(lmtest)
library(gridExtra)
library(grid)
library(forecast)
library(car)
library(gridExtra)
library(lattice)
library(stargazer)

```

# loading all the data
```{r}
df_red <- read.csv(file = 'data/winequality-red.csv', sep=';')
df_white <- read.csv(file = 'data/winequality-white.csv', sep=';')
#headers <- c('Fixed Acidity', 'Volatile Acidity','Citric Acid', 'Residual Sugars', 'chlorides', 'Free Sulfur Dioxide','Total Sulfur Dioxides', 'Density','pH', 'Sulphates', 'Alcohol' )

#df_red$Wine_Type = 'Red'
#df_white$Wine_Type = 'White'

#all_wine = rbind(df_red, df_white)
```


#Randomizing the data, before splitting to train and test
```{r}
set.seed(87323) 
df_red1 <- df_red[sample(1:nrow(df_red)), ]
set.seed(87233) 
df_white1 <- df_white[sample(1:nrow(df_white)), ]  
```

# Splitting the data to train and test for Red and White Wine
```{r}
dt_red = sort(sample(nrow(df_red1), nrow(df_red1)*.7))
train_red<-df_red[dt_red,]
test_red<-df_red[-dt_red,]

dt_white=sort(sample(nrow(df_white1), nrow(df_white1)*.7))
train_white<-df_white1[dt_white,]
test_white<-df_white[-dt_white,]

```


# Summary of White and Red Win  Dataset
```{r}

stargazer(
  df_red,
  df_white,
  type='text',title = c("Summary of Red Wine","Summary of White Wine"),summary.stat = c( "mean","median" ,"sd","min","max"), out = 'mean.html'
          
)

```




#Boxplot and Histogram Analysis for the Red and White Wine Dataset
```{r}


fig1 <- ggplot()+
geom_boxplot(data=df_red, aes(x='Red', y=fixed.acidity), color='red',fill = 'red',alpha=0.4) + 
geom_boxplot(data=df_white, aes(x='White', y=fixed.acidity), color='orange',fill = 'orange',alpha=0.4)+ scale_x_discrete(name = "")+theme(axis.line = element_line(colour = "black",size = 0.25)) +labs(title = 'Figure 11: Distribution of \n Fixed Acidity')

fig2 <- ggplot()+
geom_boxplot(data=df_red, aes(x='Red', y=volatile.acidity), color='red',fill = 'red',alpha=0.4) + 
geom_boxplot(data=df_white, aes(x='White', y=volatile.acidity), color='orange',fill = 'orange',alpha=0.4)+ scale_x_discrete(name = "")+theme(axis.line = element_line(colour = "black",size = 0.25)) +labs(title = 'Figure 12: Distribution of \n Volatile Acidity')

fig3 <- ggplot()+
geom_boxplot(data=df_red, aes(x='Red', y=citric.acid), color='red',fill = 'red',alpha=0.4) + 
geom_boxplot(data=df_white, aes(x='White', y=citric.acid), color='orange',fill = 'orange',alpha=0.4)+ scale_x_discrete(name = "")+theme(axis.line = element_line(colour = "black",size = 0.25)) +labs(title = 'Figure 13: Distribution of \n Citric Acid')

fig4 <- ggplot()+
geom_boxplot(data=df_red, aes(x='Red', y=residual.sugar), color='red',fill = 'red',alpha=0.4) + 
geom_boxplot(data=df_white, aes(x='White', y=residual.sugar), color='orange',fill = 'orange',alpha=0.4)+ scale_x_discrete(name = "")+theme(axis.line = element_line(colour = "black",size = 0.25)) +labs(title = 'Figure 14: Distribution of \n Residual Sugar')

fig5 <- ggplot()+
geom_boxplot(data=df_red, aes(x='Red', y=chlorides), color='red',fill = 'red',alpha=0.4) + 
geom_boxplot(data=df_white, aes(x='White', y=chlorides), color='orange',fill = 'orange',alpha=0.4)+ scale_x_discrete(name = "")+theme(axis.line = element_line(colour = "black",size = 0.25)) +labs(title = 'Figure 15: Distribution of Chlorides')

fig6 <- ggplot()+
geom_boxplot(data=df_red, aes(x='Red', y=free.sulfur.dioxide), color='red',fill = 'red',alpha=0.4) + 
geom_boxplot(data=df_white, aes(x='White', y=free.sulfur.dioxide), color='orange',fill = 'orange',alpha=0.4)+ scale_x_discrete(name = "")+theme(axis.line = element_line(colour = "black",size = 0.25)) +labs(title = 'Figure 16: Distribution of \n Free Sulfur Dioxide')

fig7 <- ggplot()+
geom_boxplot(data=df_red, aes(x='Red', y=total.sulfur.dioxide), color='red',fill = 'red',alpha=0.4) + 
geom_boxplot(data=df_white, aes(x='White', y=total.sulfur.dioxide), color='orange',fill = 'orange',alpha=0.4)+ scale_x_discrete(name = "")+theme(axis.line = element_line(colour = "black",size = 0.25)) +labs(title = 'Figure 17: Distribution of \n Total Sulfur Dioxide')

fig8 <- ggplot()+
geom_boxplot(data=df_red, aes(x='Red', y=density), color='red',fill = 'red',alpha=0.4) + 
geom_boxplot(data=df_white, aes(x='White', y=density), color='orange',fill = 'orange',alpha=0.4)+ scale_x_discrete(name = "")+theme(axis.line = element_line(colour = "black",size = 0.25)) +labs(title = 'Figure 18: Distribution of Density')

fig9 <- ggplot()+
geom_boxplot(data=df_red, aes(x='Red', y=pH), color='red',fill = 'red',alpha=0.4) + 
geom_boxplot(data=df_white, aes(x='White', y=pH), color='orange',fill = 'orange',alpha=0.4)+ scale_x_discrete(name = "")+theme(axis.line = element_line(colour = "black",size = 0.25)) +labs(title = 'Figure 19: Distribution of pH')


fig10 <- ggplot()+
geom_boxplot(data=df_red, aes(x='Red', y=sulphates), color='red',fill = 'red',alpha=0.4) + 
geom_boxplot(data=df_white, aes(x='White', y=sulphates), color='orange',fill = 'orange',alpha=0.4)+ scale_x_discrete(name = "")+theme(axis.line = element_line(colour = "black",size = 0.25)) +labs(title = 'Figure 20: Distribution of Sulphates')

fig11 <- ggplot()+
geom_boxplot(data=df_red, aes(x='Red', y=alcohol), color='red',fill = 'red',alpha=0.4) + 
geom_boxplot(data=df_white, aes(x='White', y=alcohol), color='orange',fill = 'orange',alpha=0.4)+ scale_x_discrete(name = "")+theme(axis.line = element_line(colour = "black",size = 0.25)) +labs(title = 'Figure 21: Distribution of Alcohol')

fig12 <- ggplot()+
geom_boxplot(data=df_red, aes(x='Red', y=quality), color='red',fill = 'red',alpha=0.4) + 
geom_boxplot(data=df_white, aes(x='White', y=quality), color='orange',fill = 'orange',alpha=0.4)+scale_x_discrete(name = "")+theme(axis.line = element_line(colour = "black",size = 0.25)) +labs(title = 'Figure 22: Distribution of Quality')


grid.arrange(fig1, fig2,fig3, fig4, ncol = 2, nrow = 2)
grid.arrange( fig5, fig6,fig7, fig8, ncol = 2, nrow = 2)
grid.arrange( fig9, fig10, fig11,fig12, ncol = 2, nrow = 2)

```





#EDA Analysis,Correleation Analysis
```{r}

all_numeric_vals_red = subset(train_red,select=-c(wine_type))
cor_tab_r <- cor(all_numeric_vals_red)
round(cor_tab_r, 2)

corrplot(cor_tab_r, type = "upper", order = "hclust", 
         tl.col = "black", tl.srt = 45, main = "Fig 13: Correlation between Variables Red Wine", mar=c(1,1,1,1))



all_numeric_vals_w = subset(train_white,select=-c(wine_type))
cor_tab_w <- cor(all_numeric_vals_w)
round(cor_tab_w, 2)

corrplot(cor_tab_w, type = "upper", order = "hclust", 
         tl.col = "black", tl.srt = 45, main = "Fig 13: Correlation between Variables White Wine", mar=c(1,1,1,1))

```



#Normality plot of the individual features for the dataset
```{r}

fig1 <- ggqqplot(df_red$fixed.acidity,title='Fixed Acidity QQ plot')
fig2 <- ggqqplot(df_red$volatile.acidity,title='Volatile Acidity \n QQ plot')
fig3 <- ggqqplot(df_red$citric.acid,title='Citric Acid QQ plot')
fig4 <- ggqqplot(df_red$residual.sugar,title='Residual Sugar \n QQ plot')
fig5 <- ggqqplot(df_red$chlorides,title='Chlorides QQ plot')
fig6 <- ggqqplot(df_red$free.sulfur.dioxide,title='Free Sulfur Dioxide \n QQ plot')
fig7 <- ggqqplot(df_red$total.sulfur.dioxide,title='Total Sulfur Dioxide \n QQ plot')
fig8 <- ggqqplot(df_red$density,title='Density QQ plot')
fig9 <- ggqqplot(df_red$pH,title='pH QQ plot')
fig10 <- ggqqplot(df_red$sulphates,title='Sulphates QQ plot')
fig11 <- ggqqplot(df_red$alcohol,title='Alcohol QQ plot')
fig12 <- ggqqplot(df_red$quality,title='Quality QQ plot')

grid.arrange(fig1, fig2,fig3, fig4, ncol = 2, nrow = 2)
grid.arrange( fig5, fig6,fig7, fig8, ncol = 2, nrow = 2)
grid.arrange( fig9, fig10, fig11,fig12, ncol = 2, nrow = 2)
```

#Variance Analysis for the X features for the  Red Wine dataset
```{r}
var_1 <- var.test(df_red$quality,df_red$fixed.acidity)
var_2 <- var.test(df_red$quality,df_red$volatile.acidity)
var_3 <- var.test(df_red$quality,df_red$citric.acid)
var_4 <- var.test(df_red$quality,df_red$residual.sugar)
var_5 <- var.test(df_red$quality,df_red$chlorides)
var_6 <- var.test(df_red$quality,df_red$free.sulfur.dioxide)
var_7 <- var.test(df_red$quality,df_red$total.sulfur.dioxide)
var_8 <- var.test(df_red$quality,df_red$density)
var_9 <- var.test(df_red$quality,df_red$pH)
var_10 <- var.test(df_red$quality,df_red$sulphates)
var_11 <- var.test(df_red$quality,df_red$alcohol)


df_var_red <- data.frame( name=  c('Fixed Acidity', 'Volatile Acidity','Citric Acid', 'Residual Sugars', 'chlorides', 'Free Sulfur Dioxide','Total Sulfur Dioxides', 'Density','pH', 'Sulphates', 'Alcohol' ),
                          var_stat =c(var_1$statistic,var_2$statistic,var_3$statistic,var_4$statistic,var_5$statistic,var_6$statistic,var_7$statistic,var_8$statistic,var_9$statistic,var_10$statistic,var_11$statistic))
df_var_red
```


#Variance Analysis for the X features for the  Red Wine dataset
```{r}
var_12 <- var.test(df_white$quality,df_white$fixed.acidity)
var_13 <- var.test(df_white$quality,df_white$volatile.acidity)
var_14 <- var.test(df_white$quality,df_white$citric.acid)
var_15 <- var.test(df_white$quality,df_white$residual.sugar)
var_16 <- var.test(df_white$quality,df_white$chlorides)
var_17 <- var.test(df_white$quality,df_white$free.sulfur.dioxide)
var_18 <- var.test(df_white$quality,df_white$total.sulfur.dioxide)
var_19 <- var.test(df_white$quality,df_white$density)
var_20 <- var.test(df_white$quality,df_white$pH)
var_21 <- var.test(df_white$quality,df_white$sulphates)
var_22 <- var.test(df_white$quality,df_white$alcohol)





df_var_white <- data.frame(name=  c('Fixed Acidity', 'Volatile Acidity','Citric Acid', 'Residual Sugars', 'chlorides', 'Free Sulfur Dioxide','Total Sulfur Dioxides', 'Density','pH', 'Sulphates', 'Alcohol' ),
                           var_stat =c(var_12$statistic,var_13$statistic,var_14$statistic,var_15$statistic,var_16$statistic,var_17$statistic,var_18$statistic,var_19$statistic,var_20$statistic,var_21$statistic,var_22$statistic))
df_var_white

library(gt)
print("Variance table for Red Wine")
gt(df_var_red,rowname_col = df_var_red$name)

print("Variance table for white Wine")
gt(df_var_white,rowname_col = df_var_white$name)

```



#Linear model for Red and White wine
```{r}
model_red_all<- lm(quality ~ fixed.acidity+volatile.acidity+citric.acid+residual.sugar+chlorides+free.sulfur.dioxide+total.sulfur.dioxide+density+pH+sulphates+alcohol+volatile.acidity , data=train_red)
model_white_all <- lm(quality ~ fixed.acidity+volatile.acidity+citric.acid+residual.sugar+chlorides+free.sulfur.dioxide+total.sulfur.dioxide+density+pH+sulphates+alcohol+volatile.acidity  , data=train_white)
summary(model_red_all)
summary(model_white_all)

library(stargazer)
stargazer(
  model_red_all,
  model_white_all,
  type='text',title = c("Summary of Red and Whites Wine"),
  header=FALSE,font.size = "huge",
  star.cutoffs=c(0.05,0.01,0.001), out="star.png"
)

```




#Linear model for Red  and White Wine Model1 Residual analysis
```{r}
print("Formal Test of Residual Normality- Red Wine")
set.seed(56423)
shapiro.test(sample(model_red_all$residuals,size=5000,replace = TRUE))
qqnorm(model_red_all$residuals)
par(mfrow=c(2,2))
plot(model_red_all)

print("Formal Test of Residual Normality- White Wine")
set.seed(56423)
shapiro.test(sample(model_white_all$residuals,size=5000,replace = TRUE))
qqnorm(model_white_all$residuals)
par(mfrow=c(2,2))
plot(model_white_all)
```

#Linear model for Red and White Wine-Model1 VIF analysis
```{r}
#create vector of VIF values
library(car)
vif_values <- vif(model_red_all)

#create horizontal bar chart to display each VIF value
par(mar=c(5,6,4,1)+2)
barplot(vif_values, main = "VIF Values Red Wine", horiz = TRUE, col = "steelblue",
        axes=TRUE, cex.names=1, las=1)

#add vertical line at 5
abline(v = 5, lwd = 3, lty = 2)

vif_values_w <- vif(model_white_all)

#create horizontal bar chart to display each VIF value
par(mar=c(5,6,4,1)+2)
barplot(vif_values_w, main = "VIF Values White Wine", horiz = TRUE, col = "steelblue",
        axes=TRUE, cex.names=1, las=1)

#add vertical line at 5
abline(v = 5, lwd = 3, lty = 2)
```



# Improved model for Red Wine
```{r}
model_red_i <- lm(quality ~ volatile.acidity+chlorides+sulphates+alcohol, data=train_red)
summary(model_red_i)
par(mfrow=c(2,2))
plot(model_red_i)

```



# Improved model for White Wine
```{r}
model_white_i <- lm(quality ~ volatile.acidity+sulphates+alcohol+residual.sugar+fixed.acidity
                    , data=train_white)
summary(model_white_i)
par(mfrow=c(2,2))
plot(model_white_i)

```

#Comparison beween the models for Red and White Wine
```{r}
library(stargazer)
stargazer(
  model_red_i,
  model_white_i,
  type='text',title = c("Summary of Red and White Wine Model"),
  header=TRUE,font.size = "huge",
  star.cutoffs=c(0.05,0.01,0.001),out = "star_1.html"
)

```



# VIF analysis for the refined model
```{r}
vif_values_r <- vif(model_red_i)

#create horizontal bar chart to display each VIF value
par(mar=c(5,6,4,1)+2)
barplot(vif_values_r, main = "VIF Values Red Wine", horiz = TRUE, col = "steelblue",
        axes=TRUE, cex.names=1, las=1)

#add vertical line at 5
abline(v = 5, lwd = 3, lty = 2)

vif_values_w_i <- vif(model_white_i)

#create horizontal bar chart to display each VIF value
par(mar=c(5,6,4,1)+2)
barplot(vif_values_w_i, main = "VIF Values White Wine", horiz = TRUE, col = "steelblue",
        axes=TRUE, cex.names=1, las=1)

#add vertical line at 5
abline(v = 5, lwd = 3, lty = 2)
```

#Anova analysis of the models
```{r}
anova(model_red_all,model_red_i ,test='F')
 anova(model_white_all,model_white_i ,test='F')


```




