---
title: "Lab 2 Lightning Ridge Proposal"
author: "Luc R, Andrew B, Vish S, Arun N"
date: '2022-03-21'
output: html_document

---


# Installing all the packages
```{r load packages, echo=FALSE, message=FALSE}

install.packages('lmtest')
install.packages('patchwork')
install.packages('sandwich')
install.packages('forecast')

library(tidyverse)
library(magrittr)
library(ggplot2)
library(patchwork)
library(sandwich)
library(lmtest)
library(gridExtra)
library(grid)
library(forecast)
library(car)

```

# loading all the data
```{r}
df_red <- read.csv(file = 'data/winequality-red.csv', sep=';')
df_white <- read.csv(file = 'data/winequality-white.csv', sep=';')
#headers <- c('Fixed Acidity', 'Volatile Acidity','Citric Acid', 'Residual Sugars', 'chlorides', 'Free Sulfur Dioxide','Total Sulfur Dioxides', 'Density','pH', 'Sulphates', 'Alcohol' )

#df_red$Wine_Type = 'Red'
#df_white$Wine_Type = 'White'

#all_wine = rbind(df_red, df_white)
```

# Data Summary for Red and White Wine-All Data
```{r}
df_red$wine_type <- "red"
df_white$wine_type <-"white"
print("Summary Red Wine")
summary(df_red)
print("Summary White Wine")
summary(df_white)
```

#Randomizing the data, before splitting to train and test
```{r}
set.seed(87323) 
df_red1 <- df_red[sample(1:nrow(df_red)), ]
set.seed(87233) 
df_white1 <- df_white[sample(1:nrow(df_white)), ]  
```

# Splitting the data to train and test for Red and White Wine
```{r}
dt_red = sort(sample(nrow(df_red1), nrow(df_red1)*.7))
train_red<-df_red[dt_red,]
test_red<-df_red[-dt_red,]

dt_white=sort(sample(nrow(df_white1), nrow(df_white1)*.7))
train_white<-df_white1[dt_white,]
test_white<-df_white[-dt_white,]

```


#Summary of all the White Wine and Red Wine Train data
```{r}
print("Summary of Red Wine Train Data")
summary(train_red)

print("Summary of White Wine Train Data")
summary(train_white)
```

```{r}
library(stargazer)
stargazer(
  df_red,
  df_white,
  type='text',title = c("Summary of Red Wine","Summary of White Wine"),summary.stat = c( "mean","median" ,"sd","min","max")
          
)


```





#EDA Analysis--> Fixed Acidity
```{r}

ggplot() + 
geom_histogram(data=df_red, aes(x=fixed.acidity , y=..density..,fill=wine_type), color='red',fill = 'red',alpha=0.25) + 
geom_histogram(data=df_white, aes(x=fixed.acidity , y=..density..,fill=wine_type), color='orange',fill = 'orange',alpha=0.25) +scale_y_continuous(name = "density") +  
       scale_x_continuous(name = "Fixed Acidity") +      
       ggtitle("Fixed Acidity Comparison") + 
       theme(axis.line = element_line(colour = "black", 
                                      size = 0.25)) +
ggplot()+
geom_boxplot(data=df_red, aes(x='Red', y=fixed.acidity), color='red',fill = 'red',alpha=0.4) + 
geom_boxplot(data=df_white, aes(x='White', y=fixed.acidity), color='orange',fill = 'orange',alpha=0.4)+ scale_x_discrete(name = "Type of Wine")+theme(axis.line = element_line(colour = "black",size = 0.25)) 

```

#EDA Analysis--> Volatile Acidity
```{r}
ggplot() + 
geom_histogram(data=df_red, aes(x=volatile.acidity , y=..density..,fill=wine_type), color='red',fill = 'red',alpha=0.25) + 
geom_histogram(data=df_white, aes(x=volatile.acidity , y=..density..,fill=wine_type), color='orange',fill = 'orange',alpha=0.25) +scale_y_continuous(name = "density") +  
       scale_x_continuous(name = "Volatile Acidity") +      
       ggtitle("Volatile Acidity Comparison") + 
       theme(axis.line = element_line(colour = "black", 
                                      size = 0.25)) +

ggplot()+
geom_boxplot(data=df_red, aes(x='Red', y=volatile.acidity), color='red',fill = 'red',alpha=0.4) + 
geom_boxplot(data=df_white, aes(x='White', y=volatile.acidity), color='orange',fill = 'orange',alpha=0.4)+ scale_x_discrete(name = "Type of Wine")+theme(axis.line = element_line(colour = "black",size = 0.25)) 
```

#EDA Analysis--> Citric Acid
```{r}
ggplot() + 
geom_histogram(data=df_red, aes(x=citric.acid  , y=..density..,fill=wine_type), color='red',fill = 'red',alpha=0.25) + 
geom_histogram(data=df_white, aes(x=citric.acid  , y=..density..,fill=wine_type), color='orange',fill = 'orange',alpha=0.25) +scale_y_continuous(name = "density") +  
       scale_x_continuous(name = "Citric Acid ") +      
       ggtitle("Citric Acid Comparison") + 
       theme(axis.line = element_line(colour = "black", 
                                      size = 0.25)) +

ggplot()+
geom_boxplot(data=df_red, aes(x='Red', y=citric.acid ), color='red',fill = 'red',alpha=0.4) + 
geom_boxplot(data=df_white, aes(x='White', y=citric.acid ), color='orange',fill = 'orange',alpha=0.4)+ scale_x_discrete(name = "Type of Wine")+theme(axis.line = element_line(colour = "black",size = 0.25)) 
```

#EDA Analysis--> Residual Sugar
```{r}

ggplot() + 
geom_histogram(data=df_red, aes(x=residual.sugar  , y=..density..,fill=wine_type), color='red',fill = 'red',alpha=0.25) + 
geom_histogram(data=df_white, aes(x=residual.sugar  , y=..density..,fill=wine_type), color='orange',fill = 'orange',alpha=0.25) +scale_y_continuous(name = "density") +  
       scale_x_continuous(name = "Residual Sugar ") +      
       ggtitle("Citric Acid Comparison") + 
       theme(axis.line = element_line(colour = "black", 
                                      size = 0.25)) +

ggplot()+
geom_boxplot(data=df_red, aes(x='Red', y=residual.sugar ), color='red',fill = 'red',alpha=0.4) + 
geom_boxplot(data=df_white, aes(x='White', y=residual.sugar ), color='orange',fill = 'orange',alpha=0.4)+ scale_x_discrete(name = "Type of Wine")+theme(axis.line = element_line(colour = "black",size = 0.25)) 
```


#EDA Analysis--> Chlorides
```{r}
ggplot() + 
geom_histogram(data=df_red, aes(x=chlorides , y=..density..,fill=wine_type), color='red',fill = 'red',alpha=0.25) + 
geom_histogram(data=df_white, aes(x=chlorides , y=..density..,fill=wine_type), color='orange',fill = 'orange',alpha=0.25) +scale_y_continuous(name = "density") +  
       scale_x_continuous(name = "Chlorides") +      
       ggtitle("Chlorides  Comparison") + 
       theme(axis.line = element_line(colour = "black", 
                                      size = 0.25)) +

ggplot()+
geom_boxplot(data=df_red, aes(x='Red', y=chlorides ), color='red',fill = 'red',alpha=0.4) + 
geom_boxplot(data=df_white, aes(x='White', y=chlorides  ), color='orange',fill = 'orange',alpha=0.4)+ scale_x_discrete(name = "Type of Wine")+theme(axis.line = element_line(colour = "black",size = 0.25)) 
```

#EDA Analysis--> Free Sulfurdioxide
```{r}
ggplot() + 
geom_histogram(data=df_red, aes(x=free.sulfur.dioxide , y=..density..,fill=wine_type), color='red',fill = 'red',alpha=0.25) + 
geom_histogram(data=df_white, aes(x=free.sulfur.dioxide  , y=..density..,fill=wine_type), color='orange',fill = 'orange',alpha=0.25) +scale_y_continuous(name = "density") +  
       scale_x_continuous(name = "Free Sulfurdioxide") +      
       ggtitle("Free Sulfurdioxide") + 
       theme(axis.line = element_line(colour = "black", 
                                      size = 0.25)) +

ggplot()+
geom_boxplot(data=df_red, aes(x='Red', y=free.sulfur.dioxide ), color='red',fill = 'red',alpha=0.4) + 
geom_boxplot(data=df_white, aes(x='White', y=free.sulfur.dioxide ), color='orange',fill = 'orange',alpha=0.4)+ scale_x_discrete(name = "Type of Wine")+theme(axis.line = element_line(colour = "black",size = 0.25)) 
```


#EDA Analysis--> Total Sulfurdioxide
```{r}
ggplot() + 
geom_histogram(data=df_red, aes(x=total.sulfur.dioxide, y=..density..,fill=wine_type), color='red',fill = 'red',alpha=0.25) + 
geom_histogram(data=df_white, aes(x=total.sulfur.dioxide , y=..density..,fill=wine_type), color='orange',fill = 'orange',alpha=0.25) +scale_y_continuous(name = "density") +  
       scale_x_continuous(name = "Total Sulfurdioxide") +      
       ggtitle("Total Sulfurdioxide") + 
       theme(axis.line = element_line(colour = "black", 
                                      size = 0.25)) +

ggplot()+
geom_boxplot(data=df_red, aes(x='Red', y=total.sulfur.dioxide ), color='red',fill = 'red',alpha=0.4) + 
geom_boxplot(data=df_white, aes(x='White', y=total.sulfur.dioxide ), color='orange',fill = 'orange',alpha=0.4)+ scale_x_discrete(name = "Type of Wine")+theme(axis.line = element_line(colour = "black",size = 0.25)) 
```


#EDA Analysis--> Density
```{r}
ggplot() + 
geom_histogram(data=df_red, aes(x= density , y=..density..,fill=wine_type), color='red',fill = 'red',alpha=0.25) + 
geom_histogram(data=df_white, aes(x= density  , y=..density..,fill=wine_type), color='orange',fill = 'orange',alpha=0.25) +scale_y_continuous(name = "density") +  
       scale_x_continuous(name = " Density ") +      
       ggtitle(" Density ") + 
       theme(axis.line = element_line(colour = "black", 
                                      size = 0.25)) +

ggplot()+
geom_boxplot(data=df_red, aes(x='Red', y= density  ), color='red',fill = 'red',alpha=0.4) + 
geom_boxplot(data=df_white, aes(x='White', y= density  ), color='orange',fill = 'orange',alpha=0.4)+ scale_x_discrete(name = "Type of Wine")+theme(axis.line = element_line(colour = "black",size = 0.25)) 
```

#EDA Analysis--> pH
```{r}
ggplot() + 
geom_histogram(data=df_red, aes(x= pH , y=..density..,fill=wine_type), color='red',fill = 'red',alpha=0.25) + 
geom_histogram(data=df_white, aes(x= pH  , y=..density..,fill=wine_type), color='orange',fill = 'orange',alpha=0.25) +scale_y_continuous(name = "pH") +  
       scale_x_continuous(name = " pH ") +      
       ggtitle(" pH Comparison") + 
       theme(axis.line = element_line(colour = "black", 
                                      size = 0.25)) +

ggplot()+
geom_boxplot(data=df_red, aes(x='Red', y= pH  ), color='red',fill = 'red',alpha=0.4) + 
geom_boxplot(data=df_white, aes(x='White', y= pH  ), color='orange',fill = 'orange',alpha=0.4)+ scale_x_discrete(name = "Type of Wine")+theme(axis.line = element_line(colour = "black",size = 0.25)) 
```

#EDA Analysis--> Sulphates
```{r}
ggplot() + 
geom_histogram(data=df_red, aes(x= sulphates , y=..density..,fill=wine_type), color='red',fill = 'red',alpha=0.25) + 
geom_histogram(data=df_white, aes(x= sulphates  , y=..density..,fill=wine_type), color='orange',fill = 'orange',alpha=0.25) +scale_y_continuous(name = "Sulphates") +  
       scale_x_continuous(name = " Sulphates ") +      
       ggtitle(" Sulphates Comparison ") + 
       theme(axis.line = element_line(colour = "black", 
                                      size = 0.25)) +

ggplot()+
geom_boxplot(data=df_red, aes(x='Red', y= sulphates  ), color='red',fill = 'red',alpha=0.4) + 
geom_boxplot(data=df_white, aes(x='White', y= sulphates ), color='orange',fill = 'orange',alpha=0.4)+ scale_x_discrete(name = "Type of Wine")+theme(axis.line = element_line(colour = "black",size = 0.25)) 
```
#EDA Analysis--> Alcohol
```{r}
ggplot() + 
geom_histogram(data=df_red, aes(x= alcohol , y=..density..,fill=wine_type), color='red',fill = 'red',alpha=0.25) + 
geom_histogram(data=df_white, aes(x= alcohol  , y=..density..,fill=wine_type), color='orange',fill = 'orange',alpha=0.25) +scale_y_continuous(name = "Sulphates") +  
       scale_x_continuous(name = " Sulphates ") +      
       ggtitle(" Alcohol Comparison ") + 
       theme(axis.line = element_line(colour = "black", 
                                      size = 0.25)) +

ggplot()+
geom_boxplot(data=df_red, aes(x='Red', y= alcohol  ), color='red',fill = 'red',alpha=0.4) + 
geom_boxplot(data=df_white, aes(x='White', y= alcohol ), color='orange',fill = 'orange',alpha=0.4)+ scale_x_discrete(name = "Type of Wine")+theme(axis.line = element_line(colour = "black",size = 0.25)) 
```

#EDA Analysis--> Quality
```{r}
ggplot() + 
geom_histogram(data=df_red, aes(x= quality , y=..density..,fill=wine_type), color='red',fill = 'red',alpha=0.25) + 
geom_histogram(data=df_white, aes(x= quality  , y=..density..,fill=wine_type), color='orange',fill = 'orange',alpha=0.25) +scale_y_continuous(name = "Quality") +  
       scale_x_continuous(name = " Quality ") +      
       ggtitle(" Quality Comparison ") + 
       theme(axis.line = element_line(colour = "black", 
                                      size = 0.25)) +

ggplot()+
geom_boxplot(data=df_red, aes(x='Red', y= quality  ), color='red',fill = 'red',alpha=0.4) + 
geom_boxplot(data=df_white, aes(x='White', y= quality ), color='orange',fill = 'orange',alpha=0.4)+ scale_x_discrete(name = "Type of Wine")+theme(axis.line = element_line(colour = "black",size = 0.25)) 
```
```{r}
library(gridExtra)
library(lattice)

fig1 <- ggplot()+
geom_boxplot(data=df_red, aes(x='Red', y=fixed.acidity), color='red',fill = 'red',alpha=0.4) + 
geom_boxplot(data=df_white, aes(x='White', y=fixed.acidity), color='orange',fill = 'orange',alpha=0.4)+ scale_x_discrete(name = "")+theme(axis.line = element_line(colour = "black",size = 0.25)) +labs(title = 'Figure 11: Distribution of \n Fixed Acidity')

fig2 <- ggplot()+
geom_boxplot(data=df_red, aes(x='Red', y=volatile.acidity), color='red',fill = 'red',alpha=0.4) + 
geom_boxplot(data=df_white, aes(x='White', y=volatile.acidity), color='orange',fill = 'orange',alpha=0.4)+ scale_x_discrete(name = "")+theme(axis.line = element_line(colour = "black",size = 0.25)) +labs(title = 'Figure 12: Distribution of \n Volatile Acidity')

fig3 <- ggplot()+
geom_boxplot(data=df_red, aes(x='Red', y=citric.acid), color='red',fill = 'red',alpha=0.4) + 
geom_boxplot(data=df_white, aes(x='White', y=citric.acid), color='orange',fill = 'orange',alpha=0.4)+ scale_x_discrete(name = "")+theme(axis.line = element_line(colour = "black",size = 0.25)) +labs(title = 'Figure 13: Distribution of \n Citric Acid')

fig4 <- ggplot()+
geom_boxplot(data=df_red, aes(x='Red', y=residual.sugar), color='red',fill = 'red',alpha=0.4) + 
geom_boxplot(data=df_white, aes(x='White', y=residual.sugar), color='orange',fill = 'orange',alpha=0.4)+ scale_x_discrete(name = "")+theme(axis.line = element_line(colour = "black",size = 0.25)) +labs(title = 'Figure 14: Distribution of \n Residual Sugar')

fig5 <- ggplot()+
geom_boxplot(data=df_red, aes(x='Red', y=chlorides), color='red',fill = 'red',alpha=0.4) + 
geom_boxplot(data=df_white, aes(x='White', y=chlorides), color='orange',fill = 'orange',alpha=0.4)+ scale_x_discrete(name = "")+theme(axis.line = element_line(colour = "black",size = 0.25)) +labs(title = 'Figure 15: Distribution of Chlorides')

fig6 <- ggplot()+
geom_boxplot(data=df_red, aes(x='Red', y=free.sulfur.dioxide), color='red',fill = 'red',alpha=0.4) + 
geom_boxplot(data=df_white, aes(x='White', y=free.sulfur.dioxide), color='orange',fill = 'orange',alpha=0.4)+ scale_x_discrete(name = "")+theme(axis.line = element_line(colour = "black",size = 0.25)) +labs(title = 'Figure 16: Distribution of \n Free Sulfur Dioxide')

fig7 <- ggplot()+
geom_boxplot(data=df_red, aes(x='Red', y=total.sulfur.dioxide), color='red',fill = 'red',alpha=0.4) + 
geom_boxplot(data=df_white, aes(x='White', y=total.sulfur.dioxide), color='orange',fill = 'orange',alpha=0.4)+ scale_x_discrete(name = "")+theme(axis.line = element_line(colour = "black",size = 0.25)) +labs(title = 'Figure 17: Distribution of \n Total Sulfur Dioxide')

fig8 <- ggplot()+
geom_boxplot(data=df_red, aes(x='Red', y=density), color='red',fill = 'red',alpha=0.4) + 
geom_boxplot(data=df_white, aes(x='White', y=density), color='orange',fill = 'orange',alpha=0.4)+ scale_x_discrete(name = "")+theme(axis.line = element_line(colour = "black",size = 0.25)) +labs(title = 'Figure 18: Distribution of Density')

fig9 <- ggplot()+
geom_boxplot(data=df_red, aes(x='Red', y=pH), color='red',fill = 'red',alpha=0.4) + 
geom_boxplot(data=df_white, aes(x='White', y=pH), color='orange',fill = 'orange',alpha=0.4)+ scale_x_discrete(name = "")+theme(axis.line = element_line(colour = "black",size = 0.25)) +labs(title = 'Figure 19: Distribution of pH')


fig10 <- ggplot()+
geom_boxplot(data=df_red, aes(x='Red', y=sulphates), color='red',fill = 'red',alpha=0.4) + 
geom_boxplot(data=df_white, aes(x='White', y=sulphates), color='orange',fill = 'orange',alpha=0.4)+ scale_x_discrete(name = "")+theme(axis.line = element_line(colour = "black",size = 0.25)) +labs(title = 'Figure 20: Distribution of Sulphates')

fig11 <- ggplot()+
geom_boxplot(data=df_red, aes(x='Red', y=alcohol), color='red',fill = 'red',alpha=0.4) + 
geom_boxplot(data=df_white, aes(x='White', y=alcohol), color='orange',fill = 'orange',alpha=0.4)+ scale_x_discrete(name = "")+theme(axis.line = element_line(colour = "black",size = 0.25)) +labs(title = 'Figure 21: Distribution of Alcohol')

fig12 <- ggplot()+
geom_boxplot(data=df_red, aes(x='Red', y=quality), color='red',fill = 'red',alpha=0.4) + 
geom_boxplot(data=df_white, aes(x='White', y=quality), color='orange',fill = 'orange',alpha=0.4)+scale_x_discrete(name = "")+theme(axis.line = element_line(colour = "black",size = 0.25)) +labs(title = 'Figure 22: Distribution of Quality')


grid.arrange(fig1, fig2,fig3, fig4, ncol = 2, nrow = 2)
grid.arrange( fig5, fig6,fig7, fig8, ncol = 2, nrow = 2)
grid.arrange( fig9, fig10, fig11,fig12, ncol = 2, nrow = 2)

```


#EDA Analysis, Linear model for Red wine-Correleation analysis
```{r}
scatterplotMatrix(~train_red$fixed.acidity+train_red$volatile.acidity+train_red$citric.acid+
                    train_red$residual.sugar+train_red$chlorides+train_red$free.sulfur.dioxide+
                    train_red$total.sulfur.dioxide+train_red$density+train_red$pH+train_red$sulphates+
                    train_red$alcohol,
                  data=NULL, plot.points= F)
```
```{r}
library(corrplot)
all_numeric_vals = subset(train_red,select=-c(wine_type))
cor_tab <- cor(all_numeric_vals)
round(cor_tab, 2)
```

```{r}
all_numeric_vals_1 = subset(train_white,select=-c(wine_type))
cor_tab_1 <- cor(all_numeric_vals_1)
round(cor_tab_1, 2)
```



#EDA Analysis,Correleation Analysis
```{r}
install.packages('corrplot')
library(corrplot)


all_numeric_vals = subset(train_red,select=-c(wine_type))
cor_tab <- cor(all_numeric_vals)
round(cor_tab, 2)

corrplot(cor_tab, type = "upper", order = "hclust", 
         tl.col = "black", tl.srt = 45, main = "Fig 13: Correlation between Variables", mar=c(1,1,1,1))
```

#EDA Analysis, Linear model for Red wine-Model1
```{r}
model_red_all<- lm(quality ~ fixed.acidity+volatile.acidity+citric.acid+residual.sugar+chlorides+free.sulfur.dioxide+total.sulfur.dioxide+density+pH+sulphates+alcohol+volatile.acidity , data=train_red)
model_white_all <- lm(quality ~ fixed.acidity+volatile.acidity+citric.acid+residual.sugar+chlorides+free.sulfur.dioxide+total.sulfur.dioxide+density+pH+sulphates+alcohol+volatile.acidity  , data=train_white)
summary(model_red_all)
summary(model_white_all)

library(stargazer)
stargazer(
  model_red_all,
  model_white_all,
  type='text',title = c("Summary of Red Wine","Summary of White Wine"),
  header=FALSE,
  star.cutoffs=c(0.05,0.01,0.001)
)

```

```{r}
install.packages('ggpubr')
library(gridExtra)
library(lattice)
library(ggpubr)
fig1 <- ggqqplot(df_red$fixed.acidity,title='Fixed Acidity QQ plot')
fig2 <- ggqqplot(df_red$volatile.acidity,title='Volatile Acidity \n QQ plot')
fig3 <- ggqqplot(df_red$citric.acid,title='Citric Acid QQ plot')
fig4 <- ggqqplot(df_red$residual.sugar,title='Residual Sugar \n QQ plot')
fig5 <- ggqqplot(df_red$chlorides,title='Chlorides QQ plot')
fig6 <- ggqqplot(df_red$free.sulfur.dioxide,title='Free Sulfur Dioxide \n QQ plot')
fig7 <- ggqqplot(df_red$total.sulfur.dioxide,title='Total Sulfur Dioxide \n QQ plot')
fig8 <- ggqqplot(df_red$density,title='Density QQ plot')
fig9 <- ggqqplot(df_red$pH,title='pH QQ plot')
fig10 <- ggqqplot(df_red$sulphates,title='Sulphates QQ plot')
fig11 <- ggqqplot(df_red$alcohol,title='Alcohol QQ plot')
fig12 <- ggqqplot(df_red$quality,title='Quality QQ plot')

grid.arrange(fig1, fig2,fig3, fig4, ncol = 2, nrow = 2)
grid.arrange( fig5, fig6,fig7, fig8, ncol = 2, nrow = 2)
grid.arrange( fig9, fig10, fig11,fig12, ncol = 2, nrow = 2)
```
```{r}
v_1 <- var.test(df_red$quality,df_red$fixed.acidity)
v_1$statistic
dim(df_red)
df_red_1 <- df_red[c(1:12)]
dim(df_red_1)


```
```{r}
cov_red <- cov(df_red_1)
library(stargazer)
stargazer(
  cov_red,colnames = NULL,
  type='text')

```


```{r}
var_1 <- var.test(df_red$quality,df_red$fixed.acidity)
var_2 <- var.test(df_red$quality,df_red$volatile.acidity)
var_3 <- var.test(df_red$quality,df_red$citric.acid)
var_4 <- var.test(df_red$quality,df_red$residual.sugar)
var_5 <- var.test(df_red$quality,df_red$chlorides)
var_6 <- var.test(df_red$quality,df_red$free.sulfur.dioxide)
var_7 <- var.test(df_red$quality,df_red$total.sulfur.dioxide)
var_8 <- var.test(df_red$quality,df_red$density)
var_9 <- var.test(df_red$quality,df_red$pH)
var_10 <- var.test(df_red$quality,df_red$sulphates)
var_11 <- var.test(df_red$quality,df_red$alcohol)


df_var_red <- data.frame( name=  c('Fixed Acidity', 'Volatile Acidity','Citric Acid', 'Residual Sugars', 'chlorides', 'Free Sulfur Dioxide','Total Sulfur Dioxides', 'Density','pH', 'Sulphates', 'Alcohol' ),
                          var_stat =c(var_1$statistic,var_2$statistic,var_3$statistic,var_4$statistic,var_5$statistic,var_6$statistic,var_7$statistic,var_8$statistic,var_9$statistic,var_10$statistic,var_11$statistic))
df_var_red

```


```{r}
var_12 <- var.test(df_white$quality,df_white$fixed.acidity)
var_13 <- var.test(df_white$quality,df_white$volatile.acidity)
var_14 <- var.test(df_white$quality,df_white$citric.acid)
var_15 <- var.test(df_white$quality,df_white$residual.sugar)
var_16 <- var.test(df_white$quality,df_white$chlorides)
var_17 <- var.test(df_white$quality,df_white$free.sulfur.dioxide)
var_18 <- var.test(df_white$quality,df_white$total.sulfur.dioxide)
var_19 <- var.test(df_white$quality,df_white$density)
var_20 <- var.test(df_white$quality,df_white$pH)
var_21 <- var.test(df_white$quality,df_white$sulphates)
var_22 <- var.test(df_white$quality,df_white$alcohol)





df_var_white <- data.frame(name=  c('Fixed Acidity', 'Volatile Acidity','Citric Acid', 'Residual Sugars', 'chlorides', 'Free Sulfur Dioxide','Total Sulfur Dioxides', 'Density','pH', 'Sulphates', 'Alcohol' ),
                           var_stat =c(var_12$statistic,var_13$statistic,var_14$statistic,var_15$statistic,var_16$statistic,var_17$statistic,var_18$statistic,var_19$statistic,var_20$statistic,var_21$statistic,var_22$statistic))
df_var_white
```

```{r}
library(gt)
print("Variance table for Red Wine")
gt(df_var_red,rowname_col = df_var_red$name)

print("Variance table for white Wine")
gt(df_var_white,rowname_col = df_var_white$name)
```


#EDA Analysis, Linear model for Red wine-Model1 Residual analysis
```{r}
print("Formal Test of Residual Normality- Red Wine")
set.seed(56423)
shapiro.test(sample(model_red_all$residuals,size=5000,replace = TRUE))
qqnorm(model_red_all$residuals)
par(mfrow=c(2,2))
plot(model_red_all)

print("Formal Test of Residual Normality- White Wine")
set.seed(56423)
shapiro.test(sample(model_white_all$residuals,size=5000,replace = TRUE))
qqnorm(model_white_all$residuals)
par(mfrow=c(2,2))
plot(model_white_all)
```

#EDA Analysis, Linear model for Red wine-Model1 VIF analysis
```{r}
#create vector of VIF values
vif_values <- vif(model_red_all)

#create horizontal bar chart to display each VIF value
barplot(vif_values, main = "VIF Values", horiz = TRUE, col = "steelblue",
        axes=TRUE, cex.names=0.5, las=1)

#add vertical line at 5
abline(v = 5, lwd = 3, lty = 2)
```

```{r}
#Model_red_2
model_red_sq <- lm(quality ~ volatile.acidity+chlorides+pH+sulphates+alcohol, data=train_red)
summary(model_red_sq)
par(mfrow=c(2,2))
plot(model_red_sq)
vif(model_red_sq)
```



#EDA Analysis, Linear model for Red wine-Model2- Interaction term

```{r}
#model_red
model_red_l <- lm(quality ~ volatile.acidity+chlorides+total.sulfur.dioxide+pH+sulphates+alcohol+sulphates*free.sulfur.dioxide  , data=train_red)
summary(model_red_l)
par(mfrow=c(2,2))
plot(model_red_l)
vif(model_red_l)
```

#EDA Analysis, Linear model for Red wine-Model2- Interaction term
```{r}
#install.packages('interactions')
library(interactions)
interact_plot(model_red_l, pred =sulphates, modx = free.sulfur.dioxide , interval = TRUE)
```
#comparison btw. models
```{r}
anova(model_red_all,model_red_sq ,test='F')
 anova(model_red_l,model_red_sq ,test='F')
 anova(model_red_all,model_red_l ,test='F')

```







#EDA Analysis, Linear model for White Wine-Model1
```{r}
model_white <- lm((quality) ~ volatile.acidity+chlorides+sulphates+alcohol+pH, data=train_white)
summary(model_white)
par(mfrow=c(2,2))
plot(model_white)
vif(model_white)
```

```{r}
model_white_i <- lm((quality) ~ volatile.acidity+chlorides+sulphates+alcohol+volatile.acidity*fixed.acidity
                    , data=train_white)
summary(model_white)
par(mfrow=c(2,2))
plot(model_white)
vif(model_white)
```

```{r}
install.packages('interactions')
library(interactions)
interact_plot(model_white_i, pred =volatile.acidity,, modx = fixed.acidity , interval = TRUE)
```



```{r}
library(stargazer)
stargazer(
  model_red_sq,
  model_white,
  type='text',
  header=FALSE,
  star.cutoffs=c(0.05,0.01,0.001)
)
```


```{r}
model_white_ov <- lm(alcohol~ volatile.acidity+chlorides+sulphates+alcohol+pH+quality, data=train_white)
summary(model_white_ov)
```

