---
title: 'W203 Final Project: Vinho Verde Wine Quality Analysis'
author: "Luc Robitaille, Andrew Bailey, Vish Sivakumar, Arun Sundrenath"
date: '2022-04-12'
output: pdf_document
---

```{r load packages, echo=FALSE, message=FALSE}

install.packages('lmtest')
install.packages('patchwork')
install.packages('sandwich')
install.packages('stargazer')
install.packages('gt')
library(tidyverse)
library(magrittr)
library(ggplot2)
library(patchwork)
library(sandwich)
library(lmtest)
library(gridExtra)
library(stargazer)
library(gt)
```

```{r Load data}
df_red <- read.csv(file = 'data/winequality-red.csv', sep=';')
df_white <- read.csv(file = 'data/winequality-white.csv', sep=';')
```


>Background: 
>The North Western portion of Portugal has a designated region from which vinho verde originates. This region is a proud producer of wine and has taken care to aggregate a large portion of data to understand what attributes are associated with the best wines. To do this a team of researchers observed over 6,000 samples of wines taking objective sensory data about the samples chemical attributes as well as having a sommelier grade the wine on a 10-point categorical scale. The test data recorded were operationalized with sensory data as well as subjective data regarding wine quality, of which the ratings were normalized by taking the median of at least 3 evaluations made by wine experts; scaling the quality between 0 (very bad) and 10 (very excellent). As a research team we hope to explore these chemical qualities and how they can be modeling indicators for high-quality wines, to help improve the product of the Vinho Verde vineyards. 
Through contemporary research surrounding the quality of wine, our team pre-posits the theory that quality of wine is the outcome of a system which relies on the balancing effects of different chemical properties. Therefore, we hypothesize that any one single variable will not be a strong indicator of a quality wine; rather that a system of indicators interacting together will support a better predictor of quality. Thus, our null hypothesis is that there is no one equation of physicochemical properties that will predict the quality of a bottle of wine.

> Research Question: 
What are the most important attributes that create a great wine?  Do red and white wines have different markers of great quality? Can we build a model which takes the chemical profile of wine and explains the overall quality of the product? 

>Explanatory Data Analysis:
To perform our Exploratory Data Analysis (EDA) the team decided to run our analysis on  the red and white wine datasets separately from each other in order to better observe the qualities of both types of wines; according to our research belief that red and white wine will have different marking indicators of quality. Furthermore, to preserve the quality of the dataset for linear model building and theory testing we separated the data to 30% EDA and 70% Linear Regression. Figures 1-12 show that there is a marked difference in the quantity of each of the wine making variables recorded in the data collection phase of this study. However, we further question, is there a parallel relationship between the quality indicators for red or white wine, or are the indicator variables different? 

```{r Tables 1 and 2 from Arun}
df_red$wine_type <- "red"
df_white$wine_type <-"white"
print("Table 1: Summary Red Wine")
summary(df_red)
print("Table 2: Summary White Wine")
summary(df_white)
```

> We were able to confirm that varieties of vinho verde wine should be treated differently and distinctly by running paired t-test across several physicochemical properties of both types.


```{r Paired T-Tested for White vs. Red}

df_wine_red_sample <- df_red[sample(nrow(df_red), 750), ]
df_wine_white_sample <- df_white[sample(nrow(df_white), 750), ]

t.test(df_wine_red_sample$fixed.acidity, df_wine_white_sample$fixed.acidity, paired=T)
t.test(df_wine_red_sample$total.sulfur.dioxide, df_wine_white_sample$total.sulfur.dioxide, paired=T)
t.test(df_wine_red_sample$density, df_wine_white_sample$density, paired=T)
t.test(df_wine_red_sample$sulphates, df_wine_white_sample$sulphates, paired=T)
```

```{r Variable Manipulation, echo= FALSE, message = FALSE}
headers <- c('Fixed_Acidity', 'Volatile_Acidity','Citric_Acid', 'Residual_Sugars', 'chlorides', 'Free_Sulfur_Dioxide','Total_Sulfur_Dioxides', 'Density','pH', 'Sulphates', 'Alcohol' )

df_red$Wine_Type = 'Red'
df_white$Wine_Type = 'White'

all_wine = rbind(df_red, df_white)

#Separate Random Samples for EDA
y <- 0.30 * nrow(df_white)
x <- 0.30 * nrow(df_red)
sample_data_white <- sample(nrow(df_white), y, replace = FALSE)
sample_data_red <- sample(nrow(df_red),x, replace = FALSE)

red_eda_df <- df_red[sample_data_red, ]
white_eda_df <- df_white[sample_data_white, ]

white_eda_df$qual_bin <- white_eda_df$quality > 6
red_eda_df$qual_bin <- red_eda_df$quality > 6

red_lm_df <- df_red[-sample_data_red,]
white_lm_df <- df_white[-sample_data_white,]

white_lm_df$qual_bin <- white_lm_df$quality > 6
red_lm_df$qual_bin <- red_lm_df$quality > 6


set.seed(87323) 
df_red1 <- df_red[sample(1:nrow(df_red)), ]
set.seed(87233) 
df_white1 <- df_white[sample(1:nrow(df_white)), ] 

dt_red = sort(sample(nrow(df_red1), nrow(df_red1)*.7))
train_red<-df_red[dt_red,]
test_red<-df_red[-dt_red,]

dt_white=sort(sample(nrow(df_white1), nrow(df_white1)*.7))
train_white<-df_white1[dt_white,]
test_white<-df_white[-dt_white,]

```

```{r EDA}
summary(all_wine)
d <- data.frame(
  type = c( white_eda_df$Wine_Type, red_eda_df$Wine_Type),
  value = c( white_eda_df$fixed.acidity, red_eda_df$fixed.acidity))

fig1 <- ggplot(data = d, aes(x=value, fill = type)) +
  geom_histogram( color="#e9ecef", alpha=0.6, position = 'identity') +
  scale_fill_manual(values=c("#404080", "#69b3a2")) + 
  labs(title = "Fig 1: Red And White \n Distribution of Fixed Acidity \n 30% Random Sample EDA")

d <- data.frame(
  type = c( white_eda_df$Wine_Type, red_eda_df$Wine_Type),
  value = c( white_eda_df$volatile.acidity, red_eda_df$volatile.acidity))

fig2 <- ggplot(data = d, aes(x=value, fill = type)) +
  geom_histogram( color="#e9ecef", alpha=0.6, position = 'identity') +
  scale_fill_manual(values=c("#404080", "#69b3a2")) + 
  labs(title = "Fig 2: Red And White \n Distribution of Volatile Acidity \n 30% Random Sample EDA")

d <- data.frame(
  type = c( white_eda_df$Wine_Type, red_eda_df$Wine_Type),
  value = c( white_eda_df$citric.acid, red_eda_df$citric.acid))

fig3 <- ggplot(data = d, aes(x=value, fill = type)) +
  geom_histogram( color="#e9ecef", alpha=0.6, position = 'identity') +
  scale_fill_manual(values=c("#404080", "#69b3a2")) + 
  labs(title = "Fig 3: Red And White \n Distribution of Citric Acid \n 30% Random Sample EDA")

d <- data.frame(
  type = c( white_eda_df$Wine_Type, red_eda_df$Wine_Type),
  value = c( white_eda_df$residual.sugar, red_eda_df$residual.sugar))

fig4 <- ggplot(data = d, aes(x=value, fill = type)) +
  geom_histogram( color="#e9ecef", alpha=0.6, position = 'identity') +
  scale_fill_manual(values=c("#404080", "#69b3a2")) + 
  labs(title = "Fig 4: Red And White \n Distribution of Residual Sugar \n 30% Random Sample EDA")

d <- data.frame(
  type = c( white_eda_df$Wine_Type, red_eda_df$Wine_Type),
  value = c( white_eda_df$chlorides, red_eda_df$chlorides))

fig5 <- ggplot(data = d, aes(x=value, fill = type)) +
  geom_histogram( color="#e9ecef", alpha=0.6, position = 'identity') +
  scale_fill_manual(values=c("#404080", "#69b3a2")) + 
  labs(title = "Fig 5: Red And White \n Distribution of Chlorides \n 30% Random Sample EDA")

d <- data.frame(
  type = c( white_eda_df$Wine_Type, red_eda_df$Wine_Type),
  value = c(white_eda_df$free.sulfur.dioxide, red_eda_df$free.sulfur.dioxide))

fig6 <- ggplot(data = d, aes(x=value, fill = type)) +
  geom_histogram( color="#e9ecef", alpha=0.6, position = 'identity') +
  scale_fill_manual(values=c("#404080", "#69b3a2")) + 
  labs(title = "Fig 6: Red And White \n Distribution of Free Sulfur Dioxide \n 30% Random Sample EDA")

d <- data.frame(
  type = c( white_eda_df$Wine_Type, red_eda_df$Wine_Type),
  value = c( white_eda_df$total.sulfur.dioxide, red_eda_df$total.sulfur.dioxide))

fig7 <- ggplot(data = d, aes(x=value, fill = type)) +
  geom_histogram( color="#e9ecef", alpha=0.6, position = 'identity') +
  scale_fill_manual(values=c("#404080", "#69b3a2")) + 
  labs(title = "Fig 7: Red And White \n Distribution of Total Sulfur Dioxide \n 30% Random Sample EDA")

d <- data.frame(
  type = c( white_eda_df$Wine_Type, red_eda_df$Wine_Type),
  value = c( white_eda_df$pH, red_eda_df$pH))

fig8 <- ggplot(data = d, aes(x=value, fill = type)) +
  geom_histogram( color="#e9ecef", alpha=0.6, position = 'identity') +
  scale_fill_manual(values=c("#404080", "#69b3a2")) + 
  labs(title = "Fig 8: Red And White \n Distribution of pH \n 30% Random Sample EDA")

d <- data.frame(
  type = c( white_eda_df$Wine_Type, red_eda_df$Wine_Type),
  value = c( white_eda_df$density, red_eda_df$density))

fig9 <- ggplot(data = d, aes(x=value, fill = type)) +
  geom_histogram( color="#e9ecef", alpha=0.6, position = 'identity') +
  scale_fill_manual(values=c("#404080", "#69b3a2")) + 
  labs(title = "Fig 9: Red And White \n Distribution of Density \n 30% Random Sample EDA")

d <- data.frame(
  type = c( white_eda_df$Wine_Type, red_eda_df$Wine_Type),
  value = c( white_eda_df$sulphates, red_eda_df$sulphates))

fig10 <- ggplot(data = d, aes(x=value, fill = type)) +
  geom_histogram( color="#e9ecef", alpha=0.6, position = 'identity') +
  scale_fill_manual(values=c("#404080", "#69b3a2")) + 
  labs(title = "Fig 10: Red And White \n Distribution of Sulphates \n 30% Random Sample EDA")

d <- data.frame(
  type = c( white_eda_df$Wine_Type, red_eda_df$Wine_Type),
  value = c( white_eda_df$alcohol, red_eda_df$alcohol))

fig11 <- ggplot(data = d, aes(x=value, fill = type)) +
  geom_histogram( color="#e9ecef", alpha=0.6, position = 'identity') +
  scale_fill_manual(values=c("#404080", "#69b3a2")) + 
  labs(title = "Fig11: Red And White \n Distribution of Alcohol \n 30% Random Sample EDA")

d <- data.frame(
  type = c( white_eda_df$Wine_Type, red_eda_df$Wine_Type),
  value = c( white_eda_df$quality, red_eda_df$quality))

fig12 <- ggplot(data = d, aes(x=value, fill = type)) +
  geom_histogram( color="#e9ecef", alpha=0.6, position = 'identity') +
  scale_fill_manual(values=c("#404080", "#69b3a2")) + 
  labs(title = "Fig 12: Red And White \n Distribution of Quality \n 30% Random Sample EDA")


grid.arrange(fig1, fig2,fig3, fig4, ncol = 2, nrow = 2)
grid.arrange(fig5, fig6, fig7, fig8, ncol = 2, nrow = 2)
grid.arrange(fig9, fig10, fig11, fig12, ncol = 2, nrow = 2)
```
> By looking at the distribution of individual physicochemical properties of the different varieties of vinho verde wine, we then confirmed with a t-test that there were significant differences between the varieties and they should be modeled differently.

>Given that the datasets we are studying are very large we will pursue a linear model under the Large Sample framework. In accordance with this framework we will model around variables which approach a normal distribution, have low correlations with other variables, and have a linear relationship with our outcome variable, quality. From our initial EDA, our dataset appears to have i.i.d underpinnings, with variables generally approaching a normal distribution. 

>Independent, identically distributed (i.i.d) data is an assumption of our dataset that will support the Central Limit Theorem (CLM), which is a statistical theorem of large datasets which will produce powerful properties for our linear regression models. An important piece of proving i.i.d is examining experimental theory and operationalization. In the case of this data set our chemical variables are all independent from another wine’s same variable measurement; e.g.  the measurement of one wine’s acidity will not influence the acidity of another wine. For our outcome variable, Quality, the score is operationalized by taking the median of three independent sommelier’s ratings (ordinal,  1-10 scale). The effect of taking a median between three independent ratings helps to decrease the effect of bias in any of the ratings which supports the i.i.d assumptions. Additionally, the i.i.d assumption requires that each data-point comes from one underlying distribution. In the case of our dataset, the data is pulled from a large sample of wines all from the Vinho Verde region of portugal. The central location of these entries provides a common distribution for all of the properties of the various wines. However, this reduces the generalizability of our results to mostly just wines in the Vinho Verde region of Portugal. 

>A final EDA check which is important to perform is the independence of individual variables relative to each other across the dataset. Figures 13 and 14 demonstrate that we have a reasonable amount of differentiation from each other to perform our linear regression analysis. 

```{r Correlation Plot for EDA}
install.packages('corrplot')
library(corrplot)


all_numeric_vals_white = subset(white_eda_df,select=-c(Wine_Type))
cor_tab_white <- cor(all_numeric_vals_white)
round(cor_tab_white, 2)

fig13 <- corrplot(cor_tab_white, type = "upper", order = "hclust", 
         tl.col = "black", tl.srt = 45, main = "Fig 13: Correlation between Variables White Wine", mar=c(1,1,1,1))

all_numeric_vals_red = subset(red_eda_df,select=-c(Wine_Type))
cor_tab_red <- cor(all_numeric_vals_red)
round(cor_tab_red, 2)

fig14 <- corrplot(cor_tab_red, type = "upper", order = "hclust", 
         tl.col = "black", tl.srt = 45, main = "Fig 14: Correlation between Variables Red Wine", mar=c(1,1,1,1))

fig13
fig14
```

>Regression Model
>Assumptions for Large Sample
Are the assumptions of the large-sample model met so that you can use an OLS regression to produce consistent estimates? 
Unique BLP Exists 

>From the analysis of the data for train data set on red wine and white wine shows unique estimates exits for the features for the individual parameters 



```{r Tables 3 and 4 from Arun}
model_red_all<- lm(quality ~ fixed.acidity+volatile.acidity+citric.acid+residual.sugar+chlorides+free.sulfur.dioxide+total.sulfur.dioxide+density+pH+sulphates+alcohol+volatile.acidity , data=train_red)
model_white_all <- lm(quality ~ fixed.acidity+volatile.acidity+citric.acid+residual.sugar+chlorides+free.sulfur.dioxide+total.sulfur.dioxide+density+pH+sulphates+alcohol+volatile.acidity  , data=train_white)
summary(model_red_all)
summary(model_white_all)

library(stargazer)
stargazer(
  model_red_all,
  model_white_all,
  type='text',title = c("Table 3: Summary of Red Wine","Table 4: Summary of White Wine"),
  header=FALSE,
  star.cutoffs=c(0.05,0.01,0.001)
)

```
>Variation in X’s 
Table below shows the variance analysis of the data set for both the Red and White wine datasets.  From the variance statistics we can conclude that the Var in most of the feature dataset for the Red and White wine dataset is > 0.2 which should be the case for satiating the assumption for the variation in the feature dataset for the large sample assumption. For both Red and White wine variance dataset, we do see a trend with the Total Sulfur Dioxide and Free Sulfur Dioxide tends to be < 0.2 for the variance statistics, this could be due to  non normality of the both the datasets and possible tight distribution of both the features. More investigation into these features is necessary to uncover more insights.


```{r}
var_1 <- var.test(df_red$quality,df_red$fixed.acidity)
var_2 <- var.test(df_red$quality,df_red$volatile.acidity)
var_3 <- var.test(df_red$quality,df_red$citric.acid)
var_4 <- var.test(df_red$quality,df_red$residual.sugar)
var_5 <- var.test(df_red$quality,df_red$chlorides)
var_6 <- var.test(df_red$quality,df_red$free.sulfur.dioxide)
var_7 <- var.test(df_red$quality,df_red$total.sulfur.dioxide)
var_8 <- var.test(df_red$quality,df_red$density)
var_9 <- var.test(df_red$quality,df_red$pH)
var_10 <- var.test(df_red$quality,df_red$sulphates)
var_11 <- var.test(df_red$quality,df_red$alcohol)


df_var_red <- data.frame( name=  c('Fixed Acidity', 'Volatile Acidity','Citric Acid', 'Residual Sugars', 'chlorides', 'Free Sulfur Dioxide','Total Sulfur Dioxides', 'Density','pH', 'Sulphates', 'Alcohol' ),
                          var_stat =c(var_1$statistic,var_2$statistic,var_3$statistic,var_4$statistic,var_5$statistic,var_6$statistic,var_7$statistic,var_8$statistic,var_9$statistic,var_10$statistic,var_11$statistic))

var_12 <- var.test(df_white$quality,df_white$fixed.acidity)
var_13 <- var.test(df_white$quality,df_white$volatile.acidity)
var_14 <- var.test(df_white$quality,df_white$citric.acid)
var_15 <- var.test(df_white$quality,df_white$residual.sugar)
var_16 <- var.test(df_white$quality,df_white$chlorides)
var_17 <- var.test(df_white$quality,df_white$free.sulfur.dioxide)
var_18 <- var.test(df_white$quality,df_white$total.sulfur.dioxide)
var_19 <- var.test(df_white$quality,df_white$density)
var_20 <- var.test(df_white$quality,df_white$pH)
var_21 <- var.test(df_white$quality,df_white$sulphates)
var_22 <- var.test(df_white$quality,df_white$alcohol)

df_var_white <- data.frame(name=  c('Fixed Acidity', 'Volatile Acidity','Citric Acid', 'Residual Sugars', 'chlorides', 'Free Sulfur Dioxide','Total Sulfur Dioxides', 'Density','pH', 'Sulphates', 'Alcohol' ),
                           var_stat =c(var_12$statistic,var_13$statistic,var_14$statistic,var_15$statistic,var_16$statistic,var_17$statistic,var_18$statistic,var_19$statistic,var_20$statistic,var_21$statistic,var_22$statistic))

print("Table 5: Variance table for Red Wine")
gt(df_var_red,rowname_col = df_var_red$name)

print("Table 6: Variance table for white Wine")
gt(df_var_white,rowname_col = df_var_white$name)
```
>Normality

>A good way to analyze the normality of the dataset is to look at the qq plot, this is a visual check for the normality of the dataset, the plot shown below shows most of the features except the pH dataset does not appear to be normally distributed, there is significant effect of the outliers on all of the features which is probably pushing the dataset to deviate from normal distribution.

```{r}

install.packages('ggpubr')
library(gridExtra)
library(lattice)
library(ggpubr)
fig1 <- ggqqplot(df_red$fixed.acidity,title='Fig 15: Fixed Acidity QQ plot')
fig2 <- ggqqplot(df_red$volatile.acidity,title='Fig 16: Volatile Acidity \n QQ plot')
fig3 <- ggqqplot(df_red$citric.acid,title='Fig 17: Citric Acid QQ plot')
fig4 <- ggqqplot(df_red$residual.sugar,title='Fig 18: Residual Sugar \n QQ plot')
fig5 <- ggqqplot(df_red$chlorides,title='Fig 19: Chlorides QQ plot')
fig6 <- ggqqplot(df_red$free.sulfur.dioxide,title='Fig 20: Free Sulfur Dioxide \n QQ plot')
fig7 <- ggqqplot(df_red$total.sulfur.dioxide,title='Fig 21: Total Sulfur Dioxide \n QQ plot')
fig8 <- ggqqplot(df_red$density,title='Fig 22: Density QQ plot')
fig9 <- ggqqplot(df_red$pH,title='Fig 23: pH QQ plot')
fig10 <- ggqqplot(df_red$sulphates,title='Fig 24: Sulphates QQ plot')
fig11 <- ggqqplot(df_red$alcohol,title='Fig 25: Alcohol QQ plot')
fig12 <- ggqqplot(df_red$quality,title='Fig 26: Quality QQ plot')

grid.arrange(fig1, fig2,fig3, fig4, ncol = 2, nrow = 2)
grid.arrange( fig5, fig6,fig7, fig8, ncol = 2, nrow = 2)
grid.arrange( fig9, fig10, fig11,fig12, ncol = 2, nrow = 2)
```

>IID
>Best way to check for IID is to look at correlation analysis between the features for the dataset, in this case we will refer to the EDA correlation of analysis of the dataset from Red wine and White Wine. Based on the analysis, the majority of the data features seem to be IID, except for pH and fixed acidity which demonstrate negative correlation. 

```{r tables7 and 8 from Arun}

```


>Omitted Variables: 

>Age:
The age of wine can be a strong indicator of quality, as it gives more time for chemicals to interact and flavor to develop. 
Year:
Wine quality can often be demarcated by the year the grapes were grown. This variable was not collected in the experiment which could be a key indicator in objective quality of the wine.
Wine Type: 
Further separation of data to the type of wine could hold information about what chemical attributes are better indicators for different types of wine. 

>Conclusion

>Creating an excellent wine is not an exact science provided just the physicochemical characteristics of the drink. It’s difficult to predict what makes an excellent wine from the major chemical variables as wine has hundreds to thousands of microbial characteristics that affect the taste, smell, and appearance of the product. Our team fails to reject the null hypothesis that: "there is no one equation of physicochemical properties that will predict the quality of a bottle of wine". However, we do believe that it is possible to predict what makes a bad bottle of wine. Through binning our data into good and bad wines it is easy to see a higher dispersion of chemical measurements on our bad bottles of wine. Thus in a future study we propose an outlier regression model that would measure for quantities of chemicals above the level accepted for good wines. 
>Practical Uses for our causal model include quality assurance for the Vinho Verde vineyards as it would be highly beneficial to understand if a wine bottle is going to be judged poorly prior to it becoming bottled. 

> Sources 

>Paulo Cortez, University of Minho, Guimarães, Portugal, http://www3.dsi.uminho.pt/pcortez
A. Cerdeira, F. Almeida, T. Matos and J. Reis, Viticulture Commission of the Vinho Verde Region(CVRVV), Porto, Portugal
@2009


