---
title: "HW week 12"
author: "w203 teaching team"
subtitle: 'w203: Statistics for Data Science'

output:
  pdf_document: default
  html_document: default
---

```{r load packages, message = FALSE}
library(tidyverse)
library(ggplot2) 
require(scales)
library(sandwich)
library(stargazer)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r source functions from project, echo = FALSE}
source('./src/load_and_clean.R')
source('./src/get_robust_se.R')
```

```{r load data} 
d <- load_and_clean(input = 'videos.txt')
```
# Regression analysis of YouTube dataset

You want to explain how much the quality of a video affects the number of views it receives on social media. In a world where people can now buy followers and likes, would such an investment increase the number of views that their content receives?  **This is a causal question.** 

You will use a dataset created by Cheng, Dale and Liu at Simon Fraser University.  It includes observations about 9618 videos shared on YouTube.  Please see [this link](http://netsg.cs.sfu.ca/youtubedata/) for details about how the data was collected.

You will use the following variables:

- `views`: the number of views by YouTube users.
- `average_rating`: This is the average of the ratings that the video received, it is a renamed feature from `rate` that is provided in the original dataset. (Notice that this is different from `cout_of_ratings` which is a count of the total number of ratings that a video has received. 
- `length:` the duration of the video in seconds.

a. Perform a brief exploratory data analysis on the data to discover patterns, outliers, or wrong data entries and summarize your findings.

```{r conduct EDA in this chunk}

summary(d)

```


```{r}

plot_length <- ggplot(data = d, aes(x = average_rating, y = length)) +
  geom_point() +
  scale_x_continuous(labels = comma) +
  scale_y_continuous(labels = comma) +
  labs(title = "Joint Distribution of Average Rating and Video Length", x = "Average Rating", y = "Video Length")

plot_log_length <- ggplot(data = d, aes(x = average_rating, y = log10(length))) +
  geom_point() +
  scale_x_continuous(labels = comma) +
  scale_y_continuous(labels = comma) +
  labs(title = "Joint Distribution of Average Rating and Log Video Length", x = "Average Rating", y = "Log 10 Video Length")

plot_length
plot_log_length

```

> 'What did you learn from your EDA? Cut this quoted text and describe your analysis in the quote block.'

b. Based on your EDA, select an appropriate variable transformation (if any) to apply to each of your three variables.  You will fit a model of the type,

$$
  f(\text{views}) = \beta_0 + \beta_1 g(\text{rate})  + \beta_3 h(\text{length})
$$ 

Where $f$, $g$ and $h$ are sensible transformations, which might include making *no* transformation. 

```{r fit a regression model here}
model <- lm(views ~ average_rating + log10(length), data=d)

 stargazer(
   model, 
   type = 'text', 
   se = list(get_robust_se(model))
   )
```



c. Using diagnostic plots, background knowledge, and statistical tests, assess all five assumptions of the CLM. When an assumption is violated, state what response you will take.  As part of this process, you should decide what transformation (if any) to apply to each variable. Iterate against your model until your satisfied that at least four of the five assumption have been reasonably addressed. 

> 1. **IID Data:** 'Replace this quote with your assessment of the IID assumption.'
> 2. **No Perfect Colinearity:** 'Replace this quote with your assessment of the no-perfect colinearity assumption.' 
> 3. **Linear Conditional Expectation:** 'Replace this quote with what you are looking for when you are assessing linear conditional expectation. Also include what you observe in your plot. In the model that you have chosen to report, do you satisfy the assumption of a linear conditional expectation? 

```{r code and plots assessing linear conditional expectation}
## remove this commented block and write code that can help you assess whether
## your model satisfied the requirement of a linear conditional expectation. 
```

> 4. **Homoskedastic Errors:** 'Replace this quote with what you are looking for when you are assessing homoskedastic errors. Also include what you observe in your plot. In the model that you have chosen to report, do you satisfy the assumption of homoskedastic errors?'

```{r code and plots assessing error variance}
## remove this commented block and write code that can help you assess whether 
## your model satisfied the requirement of homoskedastic errors
```

> 5. **Normally Distributed Errors:** 'Replace this quote with what you are looking for when you are assessing the normality of your error distribution. Also include what you observe in your plot. In the model that you have chosen to report, do you satisfy the assumption of normally distributed errors?'

```{r code and plots assessing normally distributed errors}
## remove this commented block and write code that can help you assess whether 
## your model satisfied the requirement of normally distributed errors
```